---
title: "Reproducible Research Assignment Week 2"
author: "Monique van den Broek"
date: "8 juli 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Unzip and load the data (i.e. read.csv())

```{r read data}
setwd("F:/CM-Groep/Cursussen/Data science/05_ReproducibleResearch/Week1/RepData_PeerAssessment1")
unzip("activity.zip")
data <- read.csv("activity.csv", colClasses = "character", na.strings="Not Available")
```

Process/transform the data (if necessary) into a format suitable for your analysis

```{r process data}
data$steps <- as.numeric(data$steps)
```

## Histogram of the total number of steps taken each day

Histogram of the total number of steps taken each day (ignoring the missing values in the dataset)

```{r question 2}
data_aggregated_sum <- aggregate(data$steps, list(data$date), sum , na.rm=TRUE)
hist(data_aggregated_sum$x, main="Histogram of number of steps taken per day")
```

## Mean and median number of steps taken each day

What is mean and median of total number of steps taken per day? (ignoring the missing values in the dataset)
The mean is the 4th element in the data below, the median is the 3rd element in the data below.

```{r question 3}
data_aggregated_summary <- aggregate(data$steps, list(data$date), summary , na.rm=TRUE)
data_aggregated_summary
```

## Time series plot of the average number of steps taken

```{r question 4}
data_aggregated_mean <- aggregate(data$steps, list(data$date), mean , na.rm=TRUE)
plot(as.Date(data_aggregated_mean$Group.1), data_aggregated_mean$x, type = "l", xlab= "Date", ylab= "Mean number of steps per day")
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r question 5}
data$date[[which.max(data$steps)]]
data$interval[[which.max(data$steps)]]
```

## Code to describe and show a strategy for imputing missing data

The missing values are replace by the average value for that specific interval
```{r question 6}
tapply(is.na(data$steps), data$date, FUN=sum)
# Fill with mean for that interval
data_aggregated_interval <- aggregate(data$steps, list(data$interval), mean , na.rm=TRUE)
data_aggregated_interval$interval <- data_aggregated_interval$Group.1
imputed_data <- merge(data, data_aggregated_interval, by = "interval")
for(i in 1:nrow(imputed_data)) if(is.na(imputed_data[i,]$steps)) imputed_data[i,]$steps <- imputed_data[i,]$x
```

## Histogram of the total number of steps taken each day after missing values are imputed

Calculate the total number of steps per day and the mean and median
```{r question 7}
imputed_data_sum <- aggregate(imputed_data$steps, list(imputed_data$date), sum , na.rm=TRUE)
hist(imputed_data_sum$x, main="Histogram of number of steps taken per day - imputed")
imputed_data_summary <- aggregate(imputed_data$steps, list(imputed_data$date), summary , na.rm=TRUE)
imputed_data_summary
```

The results look different. We have used the average per interval to fill the gaps, this results in more weight in the middle of the histogram. Also de medians for the days we have filled the gaps are higher than for the other days, as we assume that every hour there will be activity. Maybe it would have been better to use the median instead of the mean for filling the gaps.

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

First calculate for each date if it is week or weekend. Then take the average number of steps per interval for week/weekend and plot these results.

```{r question 8}
library(chron)
imputed_data$dow <- weekdays(as.Date(imputed_data$date),abbr = TRUE)
imputed_data$weekend = chron::is.weekend(imputed_data$date)

imputed_data_weekend <- aggregate(imputed_data$steps, list(imputed_data$weekend, imputed_data$interval), mean , na.rm=TRUE)
par(mfrow = c(1,2))
with(imputed_data_weekend[imputed_data_weekend$Group.1 == FALSE,], plot(imputed_data_weekend$Group.2, imputed_data_weekend$x, type = "l", ylab = 'Weekday Interval Means',xlab = 'Interval'))
with(imputed_data_weekend[imputed_data_weekend$Group.1 == TRUE,],  plot(imputed_data_weekend$Group.2, imputed_data_weekend$x, type = "l", ylab = 'Weekend Interval Means',xlab = 'Interval'))

```

